---
title: "Lab 3"
author: "Amelia Ritger"
date: "10/15/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

### R Markdown, data vizualization, here, dplyr::case_when()

##### Part 1: Attach packages
```{r}
library(tidyverse)
library(here) #gives R universal way to look within folders for file you're calling
library(janitor)
library(ggridges)
```

##### Part 2: Read in data from .txt
```{r}
sb_buoy <- read_table(here::here("Lab_3", "raw_data", "sb_buoy_2018.txt"),
                      na = c("99", "99.0", "99.00", "999", "999.0", "999.00")) %>% #replace missing values with various random nubers with NA
  janitor::clean_names() %>%
  dplyr::slice(-1) %>% #remove first row
  dplyr::select(number_yy:gst,atmp)
```

##### Part 3: Write a data frame to a .csv
```{r}
write_csv(sb_buoy, here::here("Lab_3", "intermediate_data", "sb_buoy.csv"))
```

##### Part 4: Wrangling to get months as an ordered factor with month abbreviations
```{r}
sb_buoy_month <-sb_buoy %>% 
  mutate(mm = as.numeric(mm)) %>% 
  mutate(month_name = month.abb[mm]) %>% # utilize month.abb function, built into base R
  mutate(month_name = fct_relevel(month_name, levels = month.abb)) # make month_name a factor, with ordered levels that match the order in month.abb
  
```

##### Part 5: Intro to {purrr} for loops
```{r}
ggplot(sb_buoy_month, aes(x=month_name, y=atmp)) +
  geom_jitter()
```

Gross! Darn R, variables should be numeric are actually still stored as characters. We want to update ALL columns currently "character" to "numeric"

```{r}
sb_buoy_num <- sb_buoy_month %>%
  purrr::modify_if(is.character, as.numeric)
```

##### Part 6: Data visualization
```{r}
ggplot(sb_buoy_num, aes(x=month_name, y=atmp)) +
  geom_jitter(aes(color=month_name),
              alpha=0.5) #make points 50% transparent

ggplot(sb_buoy_num, aes(x=month_name, y=atmp)) +
  geom_violin(aes(color=month_name,
                  fill=month_name),
              show.legend = FALSE) +
  theme_minimal() +
  facet_wrap(~month_name) #bad idea here because it makes comparisons difficult
```

Faceted graph (by month) overlayed onto population distribution:
```{r}
ggplot(sb_buoy_num, aes(x=atmp)) +
  geom_histogram(data=transform(sb_buoy_num, month_name=NULL), fill="gray90") + #transform() is base R function; plot overall population
  geom_histogram(aes(fill=month_name), show.legend = FALSE) + #create histograms color coded for each month
  facet_wrap(~month_name) + #facet by month_name
  theme_minimal()

ggsave(here::here("Lab_3", "figures", "temp_hist.png"), width=6, height=6)
```

Now, a ridgeline plot
```{r}
temp_graph <- ggplot(sb_buoy_num, aes(x=atmp, y=month_name)) +
  geom_density_ridges(aes(fill=month_name),
                      show.legend=FALSE) +
  scale_x_continuous(limits=c(5,25), #change min and max values on x axis
                     expand=c(0,0)) + #get rid of any default expansion
  theme_light()

temp_graph
ggsave(here::here("Lab_3", "figures", "temp_ridges.pdf"), width=4, height=4, dpi=150)
```

Now, let's look at windspeed and direction
```{r}
ggplot(sb_buoy_num, aes(x=wdir, y=wspd)) +
  geom_hex() +
  scale_fill_gradient(low="orange", high="blue")
```

Plot these windspeed/densities on a polar coordinate system
```{r}
ggplot(sb_buoy_num, aes(x=wdir, y=wspd)) +
  geom_density_2d(size=0.2, #make lines thinner
                  show.legend=FALSE,
                  aes(color=month_name)) +
  scale_x_continuous(breaks=c(0,90,180,270), #tell R cutoffs at 0/90/180/270
                     labels=c("N", "E", "S", "W")) + #relabel these cutoffs as compass headings
  coord_polar() +
  facet_wrap(~month_name) +
  theme_minimal()

ggsave(here::here("Lab_3", "figures", "sb_wind.png"), width=7, height=7)
```





